---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
- Extract "complex" descriptors: recurrence quantification analysis



```{r}
library(crqa)

#Read first file
d1 = read.table("pitch/Study1D0S105T1_f0.txt", header = T)

#Notes:
#Control is 0

#Extract standard descriptors
mean1 = mean(d1$f0)
std1 = sd(d1$f0)
range1 = range(d1$f0)

#Extract less standard descriptors
median1 = median(d1$f0)
iqr1 = IQR(d1$f0)
mad1 = mean(abs(d1$f0-mean(d1$f0)))
cov1 = sd(d1$f0) / mean(d1$f0)

#Extract complex descriptors
#Get optimal parameters
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
ans = optimizeParam(d1, d1, par, min.rec= 3.5, max.rec= 4.5)

#Run CRQA analysis
crqa1=crqa (d1, d1, delay=ans$delay, embed=ans$emddim, radius=ans$radius, normalize=0, rescale=0, mindiagline = 2, minvertline = 2) 
crqa1
```




2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)


```{r}

#Get a list of all the files
Files = list.files("pitch/", ".txt")
#Get a smaller list for experimentation
Files_t = sample(Files, 50)

#Make a list with the titles - later the values will be added here
d_out = c("Subj", "Study", "Diagnosis", "Trial", "Mean", "Std", "Min", "Max", "Median", "IQR", "MeanAbDif", "CoefVar", "RR", "DET", "NRLINE", "maxL", "L", "ENTR", "LAM", "TT")

#Define function for getting all those values
get_data = function(x) {
  ###Read file
  #Put the path together with the filename
  dest = paste("pitch/", x, sep="")
  #Read it
  d_t <- read.table(dest, header = T)
  
  ###Extract info from filename
  #Subject ID is the characters 10-12
  Subj = substr(x, 10, 12)
  #Diagnosis is 8th character
  Diagnosis = substr(x, 8, 8)
  #Study is 6th character
  Study = substr(x, 6, 6)
  #Trial is 14th, and possibly 15th. 
  Trial = substr(x, 14, 16)
  #Remove all non-digits in case they were there
  Trial = gsub("([0-9]+).*$", "\\1", Trial) #Remove all non-digits
  
  ###Get descriptive statistics
  #Get the mean
  mean_t = mean(d_t$f0)
  #The standard deviation
  std_t = sd(d_t$f0)
  #The minimum value
  min_t = range(d_t$f0)[1]
  #The maximum value
  max_t = range(d_t$f0)[2]
  #The median
  median_t = median(d_t$f0)
  #The inter-quartile range
  iqr_t = IQR(d_t$f0)
  #The mean absoluted deviation
  mad_t = mean(abs(d_t$f0-mean(d_t$f0)))
  #And the coefficient of variation
  cov_t = sd(d_t$f0) / mean(d_t$f0)
  
  ###Do crqa analysis
  #Get optimal parameters
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  ans = optimizeParam(d_t, d_t, par, min.rec= 3.5, max.rec= 4.5)
  
  #Run CRQA analysis
  crqa_t = crqa (d_t, d_t, delay=ans$delay, embed=ans$emddim, radius=ans$radius, normalize=0, rescale=0, mindiagline = 2, minvertline = 2) 
  
  ###Create list of results
  results = c(Subj, Study, Diagnosis, Trial, 
              mean_t, std_t, min_t, max_t, median_t, iqr_t, mad_t, cov_t, 
              crqa_t$RR, crqa_t$DET, crqa_t$NRLINE, crqa_t$maxL, crqa_t$L, crqa_t$ENTR, crqa_t$LAM, crqa_t$TT)
  
  #That list is the result of the function
           return(results)
           }


#start iteration count
n = 1

#Go through all files
for (i in Files){
#Apply the function, and add it to the list with the titles
d_out = rbind(d_out, try(get_data(i)))
  
  #Keep track of iterations
print(n)
n = n + 1
}


#Make the column names from the first row
colnames(d_out) = d_out[1, ] 
#Delete the now unnecessary first row
d_out = d_out[-1, ]  
#Make into a dataframe
d_out = data.frame(d_out)


#Rename diagnosis
library(plyr)
d_out$Diagnosis = revalue(d_out$Diagnosis, c("0" = "control", "1" = "schizo"))
```

```{r}
###### MAKE IT INTO THE DATAFRAME

#Read for further work
d_out = read.csv("Pitch_Out_Data-backupbackup.csv")
d_out$X = NULL

#Remove non-digit ID's -  the lines with errormessages
d_out = subset(d_out, grepl('^\\d+$', d_out$Subj))




#Combine ID's with Diagnosis to make unique ID's
d_out$Subj = paste(d_out$Subj, d_out$Diagnosis, sep="")


#Change types
#Make into character
d_out$Diagnosis = as.character(d_out$Diagnosis)
#Make all factors into numerical
d_out[] <- lapply(d_out, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
})
#Make into factor again
d_out$Subj = as.factor(d_out$Subj)
d_out$Diagnosis = as.factor(d_out$Diagnosis)



#### Combine with demographics data

#Read
d_demo = read.table("DemoData.txt", header = T)

#Rename for merging purposes: Diagnosis
library(plyr)
d_demo$Diagnosis = revalue(d_demo$Diagnosis, c("Control" = "control", "Schizophrenia" = "schizo"))

#Rename Subject to Subj
colnames(d_demo)[2] <- "Subj"

#Combine ID's with Diagnosis to make unique ID's
d_demo$Subj = paste(d_demo$Subj, d_demo$Diagnosis, sep="")
#Make into factor
d_demo$Subj = as.factor(d_demo$Subj)

#And combine
d_full = merge(d_out, d_demo)


#Write the datafile to csv
write.csv (d_full, file = "Pitch_Out_Data.csv")


```


3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 
3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?

4. Bonus Question: Compare effect size of diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time


```{r}



#############
library(lme4)
library(lmerTest)

#Make models for acoustic features:
model = lmer(mean ~ Diagnosis + (1|Subj), d_full)

```







